---
title: "{Fill in title}"
author: "Shilong Dai, Alex Mangiapane, <Add Your Name>"
header-includes:
- \usepackage{amsgen,amsmath,amstext,amsbsy,amsopn,amssymb,mathabx,amsthm,bm,bbm}
- \usepackage[labelsep=space]{caption}
output:
  pdf_document: default
  word_document: default
  html_document: default
subtitle: $\textbf{Machine Learning, Spring 2021}$
---

# Abstract
*{Fill in your abstract}*

# Introduction
*{Fill in your introduction}*

# Setup
```{r}
set.seed(315)
# Load necessary packages
library(dslabs)
library(ggplot2)
```


# Loading and Preprocessing the Data
1. Begin by loading the dataset you've chosen. Depending on which dataset you've chosen, it may be helpful to consider a subset of the data for your analyses. Give an explanation as to why you did or did not subset your data.
```{r}
minst_data <- read_mnist()
minst.train_raw <- minst_data$train
minst.test_raw <- minst_data$test

```

2. Next, provide some basic information about the data. What are the samples and what are the variables? How many samples are there? How many variables are there? How many of those variables are numerical vs categorical? How many of the numerical variables are discrete? Discuss any other information about the data that you think is relevant.
```{r}
print("Training Sample Size")
length(minst.train_raw$labels)
print("Training Image Size")
length(minst.train_raw$images[1, ])
print("Training Number 1")
minst.train_raw$labels[1]
print("Training Number Range")
sort(unique(minst.train_raw$labels))
print("Test Sample Size")
length(minst.test_raw$labels)
print("Test Image Size")
length(minst.test_raw$images[1, ])
print("Test Number 1")
minst.test_raw$labels[1]
print("Test Number Range")
sort(unique(minst.test_raw$labels))
```

The data consists of a set of training and test samples. For both the training and test samples, the sample consists of a sequence of images in greyscale and labels. Each image is flattened from 2D to a 1D sequence of values between 0 and 255, which indicates the blackness of the pixel. After flattening, there are 784  Each label is a number corresponding to the actual handwritting digit of the associated image.

There are a total of 60000 training samples, and 10000 test samples. Since all representations are either digital or labels indicating the written digit, both the greyscale value and the digit label are discrete. Each greyscale value corresponds to an pixel and thus is natural numbers on [0, 255]. The label of the numbers themselves are categorical variables between 0 and 9, representing the nine digits.

3. In some scenarios, it may be helpful to transform or preprocess your data in some way before embarking on your analysis. For example, one may wish to center and standardize each numerical variable before performing certain regression analyses. On the other hand, one may wish to group some factors of a categorical variable together. Perform any preprocessing that you feel is appropriate and discuss the motivation for this choice. If you decide not to, explain why. Note that it may be helpful to omit this step and return at a later point once you have explored the data a bit.
```{r}
preprocess_minst <- function (images, labels) {
  image_col_names = c("digit")
  for (i in 1:length(images[1, ])) {
    image_col_names = c(image_col_names, sprintf("p_%d", i))
  }

  result <- as.data.frame(cbind(labels, images))
  colnames(result) <- image_col_names
  return(result)
}

split_minst <- function (minst, digit, train_amount = 2000, test_amount = 1000) {
  selected_digit <- minst[minst[["digit"]] == digit, ]
  train_rows <- sample(nrow(selected_digit), train_amount)
  train_sample <- minst[train_rows, ]
  remainder <- minst[-train_rows, ]
  test_rows <- sample(nrow(remainder), test_amount)
  test_sample <- remainder[test_rows, ]
  return(list(train = train_sample, test = test_sample))
}

minst.train <- preprocess_minst(minst.train_raw$images, minst.train_raw$labels)
minst.test <- preprocess_minst(minst.test_raw$images, minst.test_raw$labels)
minst.combined <- rbind(minst.train, minst.test)

minst_final.test <- data.frame()
minst_final.train <- data.frame()
for(i in 0:9) {
  sampled <- split_minst(minst.combined, i)
  minst_final.test <- rbind(minst_final.test, sampled$test)
  minst_final.train <- rbind(minst_final.train, sampled$train)
}
minst_final.together <- rbind(minst_final.train, minst_final.test)


head(minst_final.train)
head(minst_final.test)
```

The images and labels are combined and matched together into the sample and test dataframes. The rows in each dataframe represents the pixels of an image with its label. This is done in order to have the image and the label variable together for easier processing.

# Exploratory Data Analysis
4. Provide a statistical summary of the data as well as a summary of the data in words. Note that it is not necessary to list every summary statistic for every variable. Rather, try to give the reader a sense of what kinds of values the variables take on. And don't forget any categorical variables!
```{r, eval=FALSE}

Mode <- function(x) {
  ux <- unique(x)
  return(ux[which.max(tabulate(match(x, ux)))])
}

image_only_minst <- minst_final.together[, -c(1)]
flattened_images <- as.vector(t(image_only_minst))

center_pixels <- as.vector(t(minst_final.together["p_392"]))
edge_pixels <- as.vector(t(minst_final.together['p_1']))
rand_pixels <- as.vector(t(minst_final.together['p_457']))

flattened_sum <- summary(flattened_images)
center_sum <- summary(center_pixels)
corner_sum <- summary(edge_pixels)
rand_sum <- summary(rand_pixels)

sd_flattened <- sd(flattened_images)
sd_center <- sd(center_pixels)
sd_corner <- sd(edge_pixels)
sd_rand <- sd(rand_pixels)
sd_sum <- c(sd_flattened, sd_center, sd_corner, sd_rand)

mode_flattened <- Mode(flattened_images)
mode_center <- Mode(center_pixels)
mode_corner <- Mode(edge_pixels)
mode_rand <- Mode(rand_pixels)
mode_sum <- c(mode_flattened, mode_center, mode_corner, mode_rand)

sum_table <- rbind(flattened_sum, center_sum, corner_sum, rand_sum)
sum_table <- cbind(sum_table, sd_sum, mode_sum)

colnames(sum_table) <- c(colnames(sum_table)[1:6], "StDev", "Mode")
rownames(sum_table) <- c("All", "Center", "Corner", "P. 457")
print(as.data.frame(sum_table))
```

Rather than give a summary for each of the 784 pixels in each image, instead we gave a summary of four important categories: the entire set of pixel values, the center pixel (Pixel 384), a corner pixel (Pixel 1), and a random pixel (Pixel 457). It comes as no surprise that the center pixels are much darker on average than corner pixels, given how digits tend to be centerred in the middle of the image. In fact, given that the corner pixel has a max of 0, no writing was in the corner of any image. Also note that the centered pixel, as expected has a higher standard deviation than the corner pixel. Our randomly chosen pixel happened to be a darker than average pixel, as clearly displayed with a mean of 67.7 darkness, and with a higher than average standard deviation. Also note that the mode of each of thes categories is 0, meaning even in the center pixels, the most likely scenario is no writing in that pixel.

5. Pick three variables in your dataset and visualize the distribution of each in separate plots. Remember to properly title and label each plot. Compare them and comment on what you see. Do any of the distributions appear similar or different? In what ways?
```{r}
ggplot(data = minst_final.together, aes(x = p_457)) + geom_histogram() + ggtitle("Distribution of Pixel 457") + xlab("Pixel Values") + ylab("Frequency")
ggplot(data = minst_final.together, aes(x = p_1)) + geom_histogram(binwidth = 1) + ggtitle("Distribution of Pixel 1") + xlab("Pixel Values") + ylab("Frequency")
ggplot(data = minst_final.together, aes(x = p_536)) + geom_histogram() + ggtitle("Distribution of Pixel 536") + xlab("Pixel Values") + ylab("Frequency")
```
As is clear in the histograms, we can clearly see that in pixel one, every image had a value of 0. Our other two chosene pixels, Pixel 457 and Pixel 536, had more interesting distributions. Pixel 457 still had mostly 0's, but also many 255's, with some much smaller frequency values spread out in between. Pixel 536 also had mostly 0's, with a very small amount of 255's, and very low frequency values in between. As such, Pixel 536 resembled Pixel 1 fairly closely, with almost entirely 0's, whereas Pixel 457 had a slightly greater range of values.

6. Pick two numerical variables in your dataset. What is their correlation? Visualize their joint behavior in a single plot. Does this suggest that there is a strong relationship between the two?
```{r}
cor(minst_final.together["p_455"], minst_final.together["p_456"])
plot(minst_final.together$p_455, minst_final.together$p_456, main = "Plot of Pixel 455 Values Vs. Pixel 456 Values", xlab = "Pixel 455 Values", ylab = "Pixel 456 Values")
```
Here, we arbitrarily chose to find the correlation between two pixels right next to each other (Pixels 455 and 456), expecting a reasonably strong correlation because of their proximity. We calculated the correlation between these two pixel values to be 0.781128, a strong correlation. As can be seen in the plot, there are many data points clustered in the top corners of the plot, with the data becoming increasingly sparse toward the middle. Thus, the plot does suggest a reasonable amount of correlation betweene the variables. 

7. Pick two numerical variables in your dataset. Perform a hypothesis test to check whether their means are equal. Discuss what test you performed and why. What is the result of the test? Note that such a test won't make sense if you have centered your data so you should apply this to your data without any centering.
```{r}
mean(minst_final.together$p_455)
mean(minst_final.together$p_456)
wilcox.test(minst_final.together$p_455, minst_final.together$p_456)
```
We can initially see that the means of the pixel darkness in Pixel 456 and Pixel 455 are quite different, so to determine if this is statistically significant, we use wilcox.test (since the data is paired and is not normally distributed, so we cannot just use t.test). Here, our p-value is extremely close to 0, so we can confidently reject our null hypothesis and conclude that these to variables have unequal means. Note: this test is known as the Wilcoxon Signed Rank Test.

8. In a single plot, visualize the relationship between the means or medians and the standard deviations of each variable. Do you notice any outliers? 
```{r, eval=FALSE}
means <- apply(minst_final.together, 2, mean)
sds <- apply(minst_final.together, 2, sd)
plot(means, sds, main = "Plot of Means vs. Standard Deviations", xlab = "Means", ylab = "Standard Deviations")
```
Looking at the graph, there does appear to be one serious outlier near the bottom left, but otherwise the points all follow a very clear trend.

# Principal Components Analysis
9. Next, we would like to reduce the dimension of the data. Explain what the "dimension" of a dataset is. Discuss why one would want to reduce it. Provide an intuitive explanation of how PCA can achieve this.

The dimension of a dataset is the number of variables it is tracking. In our example, we have 784 pixels of which we are tracking the darkness. This is a lot of variables to keep track of, and may often cause unnecessary expenses or time-wasting in collecting and analyzing so many different data components. As such, it is often desirable to reeduce the data into a more manageable number of components in order to save time and money, particularly if the dataset can be reduced without losing much information conveyed in the data. For instance, with our dataset, it would be highly beneficial if we could reduce the 784 pixels we track to less than 100, or even fewer, and still be able to accurately determine the value of the hand-drawn number. PCA achieves this by analyzing how much each component variable contributes to predicting the label, and then removing the components that contribute very little to the overall prediction. In our example, we may likely end up finding that the pixels around the border of our image, given that they are usually completely white, do not contribute to predicting the label, and thus we may reasonable stop tracking these pixels. 

10. Run PCA on the dataset. Decide whether to set *scale = TRUE* or *scale = FALSE* and explain your choice. Provide a numerical summary of the first 5 PCs. Note that you should leave any categorical variables out of this analysis.
```{r}
minst_final.together.numerical <- minst_final.together[, -c(1)]
minst_pcs <- prcomp(minst_final.together.numerical)
minst_pcs_sum <- summary(minst_pcs)
print(as.data.frame(minst_pcs_sum$importance[, 1:5]))
```

11. Plot a screeplot of the PCs. How many principal components are required to explain at least 80% of the variation in the data? Based on this and the plot, does it seem like PCA has done a good job in reducing the dimensionality of the data?
```{r, eval=FALSE}
screeplot(minst_pcs, npcs = 20, type = "lines", main = "Variance explained by PC")
cum_pcs_filtered <- minst_pcs_sum$importance[3, abs(minst_pcs_sum$importance[3, ] - 0.8) < 0.01]
cum_pcs_filtered <- as.data.frame(as.list(cum_pcs_filtered))
rownames(cum_pcs_filtered) <- c("Cum. Variance Explained")
print(cum_pcs_filtered)
```

12. Visualize the distributions of the first 3 PCs in separate plots and comment. Are there any apparent clusters? Do any of the first 3 PCs appear most helpful for separating the data?
```{r, eval = FALSE}
minst_pcs_df.3 <- data.frame(
  IDX = rownames(minst_final.together.numerical),
  PC1 = minst_pcs$x[, 1],
  PC2 = minst_pcs$x[, 2],
  PC3 = minst_pcs$x[, 3]
) 
minst_pcs_df.3 <- cbind(minst_final.together["digit"], minst_pcs_df.3)
minst_pcs_df.3[, "digit"] <- factor(minst_pcs_df.3[, "digit"])

# TODO: figure out how to combine the three graph in one frame. Alternatively, use the default plot, but it is a pain to deal with in terms of adding legend for color

for(i in 1:3) {
  col <- paste0("PC", i)
  print(ggplot(data = minst_pcs_df.3, aes(color = digit)) + aes_string(x = "IDX", y = col) + geom_point() + ggtitle(paste0("PC", i, " with Digit")) + xlab("Sample Number") + ylab("PC Score"))
}

```


13. Plot the first 3 PCs against one another (see the Biplots section of Computing Assignment 3). Are there any apparent clusters?
```{r, eval = FALSE}
pairs(minst_pcs_df.3[, 3:5], col = minst_pcs_df.3$digit, oma=c(3,3,3,15))
par(xpd = TRUE)

# TODO: might need to tweek the size of the legend, and add label
legend("bottomright", fill = unique(minst_pcs_df.3$digit), legend = c( levels(minst_pcs_df.3$digit)))
```

# Clustering
14. Next we would like to perform a cluster analysis on the data. Explain intuitively what that means. What exactly is a cluster? What are the objects being clustered? Why might this be helpful for this dataset?

15. Apply one of the clustering algorithms you learned about in class to your data. Depending on the algorithm you use, you may need to specify a cutpoint to obtain a single cluster label for each sample. Note also that depending on the dimension of your dataset, it may be preferable to apply the clustering algorithm to the first few PCs of your data rather than the data itself. Elaborate upon your choice.
```{r, eval = FALSE}
# Apply the clustering algorithm.
```

16. What did the clustering algorithm find? Are the clusters relatively homogeneous or heterogeneous? Summarize the results.
```{r, eval = FALSE}
# Summarize results of clustering.
```

17. What is the within-cluster sum of squares for the clusters you found? How does this compare to the total sum of squares?
```{r, eval = FALSE}
# Compute sums of squares.
```

18. Plot the first 3 PCs against one another again but include the cluster label for each point. Do any patterns emerge? Comment on what you see.
```{r, eval = FALSE}
# Plot PCs with clusters.
```

# Classification
19. Next we would like to perform classification on the data. Explain intuitively what that means. What does it mean to classify the data? What are the objects being classified? Why might this be of interest for this dataset?

20. Identify a categorical variable in the data to correspond to the class of each sample in the dataset. In particular, is there a variable that might be interesting to predict from the other variables?

21. Construct a table that describes how many samples in each class fall into each cluster. How well was your previous cluster analysis able to identify the classes you chose? Are any of the clusters comprised mostly by one class?
```{r, eval=FALSE}
# Construct table.
```

22. Randomly break your dataset into a training set (roughly 80% of the data) and a test set (roughly 20% of the data). What is the point of doing this before doing classification?
```{r, eval = FALSE}
# Break into train and test sets.
```

23. Using the class labels based on the categorical variable you chose earlier, apply a classification method to the training set. Be sure to specify any free parameters that were chosen. What is the accuracy of the fitted model on the training set?
```{r, eval = FALSE}
# Run classification algorithm.
# Compute training set accuracy.
```

24. What is the accuracy of the fitted model on the test set? How does this compare to the accuracy on the training set? Does this make sense? Explain why or why not.
```{r, eval = FALSE}
# Compute accuracy on test set.
```

25. Construct a confusion matrix based on the results of classification. Were there any classes on which your algorithm performed better than others? What about worse than others? Does this make sense based on the nature of the data?
```{r, eval = FALSE}
# Find classes where the algorithm performed best / worst.
```


# Conclusion
*{Fill in your conclusion}*

